### 2015-06-03 review1/response.txt

Dear editors, dear referee,

thank you for your valuable comments concerning our manuscript. We have addressed all of them and corrected the text accordingly. Please find detailed answers below.

Best regards,
Jakub Urban

> Reviewer #2: The manuscript is well organized and written in general. I found some misprints and other small inconsistencies to check for the final version. I have also some comments.
> Misprints.
> 1. Introduction. 
> now: … which predicts PF coils currents from a give plasma boundary and profiles.
> to be: … which predicts PF coils currents from a given plasma boundary and profiles.
> 3.2 Statistical analysis
> now: For shot 6962, equilibria with TS pressure profiles and with linear p' and FF'.
> to be: For shot 6962, equilibria with TS pressure profiles and with linear p' and FF' is used.
> 3.2 Statistical analysis
> now: We can observer that the…
> to be: We can observe that the…

All the misprints have been corrected.

> Inconsistencies.
> 
> 2. Verification and validation procedure.
> When defining verification and validation procedure you specify 4 steps. Then you expain them, but in explanation you change order:
> Procedure definition:
> 4. Reconstruct FREEBIE equilibria using EFIT++ and VacTH with various parameters and artificial input noise.
> Later in text:
> In the optional third step, an artificial random noise is added to the calculated values of Ip, magnetic probes and flux loops.

We have changed the text of chapter 2:
" ...
FREEBIE inputs can be modified in the optional third step. In the following analysis, we particularly use realistic pressure profiles, estimated by Thomson scattering diagnostics.

The final fourth step consists of reconstructing the equilibria form synthetic FREEBIE data using EFIT++ and VacTH. An artificial random noise is added to the calculated values of $I_\mathrm{p}$, magnetic probes and flux loops. In particular, for a given noise level $\epsilon$, $\tilde X = \left( {1 + U\left( { - \epsilon, \epsilon} \right)^\mathrm{T}} \right)X$, where $X$ is a row vector of the synthetic diagnostics data and $U\left( { - \epsilon, \epsilon} \right)$ is a random vector of the same shape as $X$ with a uniform distribution on $\left( { - \epsilon, \epsilon} \right)$. The reconstructions are then compared to the original equilibrium, focusing on global parameters and geometry. Scans are performed over noise levels ($\epsilon$) and selected code parameters: $p'$ and $FF'$ polynomial degrees in EFIT++ ($n_{p'}$, $n_{FF'}$) and the number of harmonics ($n_P$, $n_Q$) in VacTH. The following quantities are used for the comparison.
... "

> 3. Results.
> You introduce different shapes as circular elongated and diverted. Then in 3.1 you are using term 'eliptical' several times, that can refer to both elongated and diverted plasmas. Could you specify which plasmas you refer to in each case.

We now use 'elliptical' throughout the paper when referring to elongated, limited plasmas. In the beginning of chapter 3, we now explicitly say: "These cases include circular, elliptical (limited) and diverted plasmas with different currents." We believe that this avoids any confusion since diverted plasmas are always designated explicitly.

> Comments.
> 
> 1. When talking about ITM CPOs, I think you need reference to explain what is it. It could be for example:
> G. Manduchi et al, Fusion Eng. Design 83 (2008) 462

The reference has been added.

> 2. Annotation to Fig.2. Last sentence is not clear for me, what do you mean? Have you artificially scattered zero error data? Could you explain this?

We simply scattered the x position of the data points with zero input error in order to suppress their overlap. We have removed this sentence from the caption as it should be clear enough that the error of these data points is zero. Based on your comment, the sentence was a bit confusing.

> 3. Fig.2. I propose to change color for case s3 from purple to (for example) green to distinguish results (it is hard to do now).

The colour has been changed.

> 4. Conclusions. When you are saying that VacTH is a promising tool for a real-feedback control, on what reasoning are you relying? In the text it is shown that VacTH is able to reconstruct plasma surface with 'reasonable' accuracy. I think in order to conclude about it's applicability for real-time control it should be verified against another real-time tools. My conclusion from the results shown is that further improvements in the code are needed to make it more robust (that is required for real-time feedback control).

The conclusion we draw in the paper is that VacTH is promising for real-time shape control. In out opinion, the accuracy of VacTH shape reconstruction shall be reasonable for real time plasma position/shape control. However, we agree that the robustness (and the accuracy as well) should be improved. There are actually new results for VacTH parameters optimizations. We do not think that VacTH must necessarily be compared to other real-time shape reconstruction tools providing that it shows enough accuracy, robustness and speed required by a particular feed-back algorithm. Since such an algorithm does not exist yet on COMPASS, we do not have absolute measures.
We have extended the last sentence of the conclusions: "We conclude that VacTH is a promising tool pertinent for a real-time feedback control of the plasma shape, although its robustness and accuracy should be improved."

### 2015-06-03 Conclusions.tex
submitted version: * no subdirectories * removed comments * added cover letter and research highlights
Response to review (first version, TO BE CHECKED)

Two new codes---FREEBIE and VacTH---have been successfully set up on COMPASS, which enabled to perform an extensive cross-benchmarking and validation of free-boundary equilibrium tools. We show that FREEBIE can predict equilibria that are consistent with EFIT++ reconstructions from experimental data. FREEBIE model equilibria, either with linear $p'$ and $FF'$ profiles or with pressure profiles from Thomson scattering diagnostic, then served to assess the credibility of EFIT++ reconstructions. 
% This has not been up to now possible.

We show that magnetic reconstruction EFIT++ with linear $p'$ and $FF'$ features a relatively good accuracy of 1 -- 2 cm in the plasma shape reconstruction but introduces a systematic error in internal plasma parameters, such as $W$, $l_{\mathrm i}$, $\beta_{\mathrm p}$ or $q_0$. The reconstruction properties can be significantly improved by using quadratic $p'$ for diverted plasmas, which removes the systematic error and also improves the LCFS reconstruction. EFIT++ converges in 100~\% cases in this regime.

Optimum parameters for VacTH have been estimated. In particular, the optimum number of harmonics is 4 otherwise VacTH fails to converge in many cases, even without any input error. 16 flux loops and only 8 magnetic must be used as VacTH input. With less flux loops or more magnetic probes the code performs significantly worse. We conclude that VacTH is a promising tool pertinent for a real-time feedback control of the plasma shape.

>>>>

Two new codes---FREEBIE and VacTH---have been successfully set up on COMPASS, which enabled to perform an extensive cross-benchmarking and validation of free-boundary equilibrium tools. We show that FREEBIE can predict equilibria that are consistent with EFIT++ reconstructions from experimental data. FREEBIE model equilibria, either with linear $p'$ and $FF'$ profiles or with pressure profiles from Thomson scattering diagnostic, then served to assess the credibility of EFIT++ reconstructions. 
Magnetic EFIT++ reconstruction with linear $p'$ and $FF'$ features a relatively good accuracy of 1 -- 2 cm in the plasma shape but introduces systematic errors both in the shape and in internal plasma parameters, such as $W$, $l_{\mathrm i}$, $\beta_{\mathrm p}$ or $q_0$. The reconstruction properties can be significantly improved by using quadratic $p'$ for elliptical and divertor plasmas, which removes the systematic error and also improves the LCFS reconstruction. EFIT++ converges in 100~\% cases in this regime.
Optimum parameters for VacTH have been estimated. In particular, the optimum number of harmonics is 4 otherwise VacTH fails to converge in many cases, even without any input error. 16 flux loops and only 8 magnetic probes must be used as VacTH input. With less flux loops or more magnetic probes the code performs significantly worse. We conclude that VacTH is a promising tool pertinent for a real-time feedback control of the plasma shape, although its robustness and accuracy should be improved.

### 2015-06-03 Procedure.tex
Response to review (first version, TO BE CHECKED)

It should be noted here that to set up a free-boundary equilibrium code, a rather complete machine description is necessary (in particular, the PF coils geometry and circuits, limiter, vessel and other passive PF elements and magnetic diagnostics configuration). We adopted the description that was already available for EFIT++ and transformed it to ITM CPO's (Integrated Tokamak Modelling Consistent Physical Objects) structures, which are subsequently either used directly in FREEBIE or converted to VacTH specific input format.

@@

In the optional third step, an artificial random noise is added to the calculated values of $I_\mathrm{p}$, magnetic probes and flux loops. In particular, for a given noise level $\epsilon$, $\tilde X = \left( {1 + U\left( { - \epsilon, \epsilon} \right)^\mathrm{T}} \right)X$, where $X$ is a row vector of the synthetic diagnostics data and $U\left( { - \epsilon, \epsilon} \right)$ is a random vector of the same shape as $X$ with a uniform distribution on $\left( { - \epsilon, \epsilon} \right)$.

The final fourth step consists of reconstructing the equilibria form synthetic FREEBIE data using EFIT++ and VacTH. The reconstructions are then compared to the original equilibrium, focusing on global parameters and geometry. Scans are performed over noise levels ($\epsilon$) and selected code parameters: $p'$ and $FF'$ polynomial degrees in EFIT++ ($n_{p'}$, $n_{FF'}$) and the number of harmonics ($n_P$, $n_Q$) in VacTH. The following quantities are used for the comparison.
\begin{table}[!h]
    \begin{tabular}{ll}
    $R_{\mathrm ax}$, $Z_{\mathrm ax}$ & $R,Z$ coordinates of the magnetic axis \\
    $R_{\mathrm in}$, $R_{\mathrm out}$ & inner/outer $R$ coordinate of LCFS at $Z=Z_{\mathrm ax}$ \\
    $Z_{\mathrm min}$, $Z_{\mathrm max}$ & minimum/maximum $Z$ coordinate of LCFS \\
    $I_{\rm{p}}$ & plasma current \\
    $\kappa  = \frac{\left( {{Z_{{\rm{max}}}} - {Z_{{\rm{min}}}}} \right) } { \left( {{R_{{\rm{out}}}} - {R_{{\rm{in}}}}} \right) }$          & elongation   \\
    ${l_{\rm{i}}} = {{\bar B_{\rm{p}}^2}} / {{B_{\rm{a}}^2}}$          & normalized internal inductance   \\
    ${\beta _{\rm{p}}} = {{2{\mu _0}\bar p}} / {{B_{\rm{a}}^2}}$          & poloidal beta   \\
    $ W = \int_0^V {\frac{3}{2}p{\rm{d}}V'}$ & stored plasma energy \\
    $ q_0$, $q_{95}$ & safety factor at $\bar \psi = 0,\ 0.95$ \\
    \end{tabular}
\end{table}

>>>>

It should be noted here that to set up a free-boundary equilibrium code, a rather complete machine description is necessary (in particular, the PF coils geometry and circuits, limiter, vessel and other passive PF elements and magnetic diagnostics configuration). We adopted the description that was already available for EFIT++ and transformed it to ITM CPO's (Integrated Tokamak Modelling Consistent Physical Objects \cite{Manduchi2008}) structures, which are subsequently either used directly in FREEBIE or converted to VacTH specific input format.

@@

FREEBIE inputs can be modified in the optional third step. In the following analysis, we particularly use realistic pressure profiles, estimated by Thomson scattering diagnostics.

The final fourth step consists of reconstructing the equilibria form synthetic FREEBIE data using EFIT++ and VacTH. 
An artificial random noise is added to the calculated values of $I_\mathrm{p}$, magnetic probes and flux loops. In particular, for a given noise level $\epsilon$, $\tilde X = \left( {1 + U\left( { - \epsilon, \epsilon} \right)^\mathrm{T}} \right)X$, where $X$ is a row vector of the synthetic diagnostics data and $U\left( { - \epsilon, \epsilon} \right)$ is a random vector of the same shape as $X$ with a uniform distribution on $\left( { - \epsilon, \epsilon} \right)$.
The reconstructions are then compared to the original equilibrium, focusing on global parameters and geometry. Scans are performed over noise levels ($\epsilon$) and selected code parameters: $p'$ and $FF'$ polynomial degrees in EFIT++ ($n_{p'}$, $n_{FF'}$) and the number of harmonics ($n_P$, $n_Q$) in VacTH. The following quantities are used for the comparison.
\begin{table}[!h]
    \begin{tabular}{ll}
    $R_{\mathrm ax}$, $Z_{\mathrm ax}$ & $R,Z$ coordinates of the magnetic axis \\
    $R_{\mathrm in}$, $R_{\mathrm out}$ & inner/outer $R$ coordinate of LCFS at $Z=Z_{\mathrm ax}$ \\
    $Z_{\mathrm min}$, $Z_{\mathrm max}$ & minimum/maximum $Z$ coordinate of LCFS \\
    $I_{\rm{p}}$ & plasma current \\
    $\kappa  = \frac{\left( {{Z_{{\rm{max}}}} - {Z_{{\rm{min}}}}} \right) } { \left( {{R_{{\rm{out}}}} - {R_{{\rm{in}}}}} \right) }$          & elongation   \\
    ${l_{\rm{i}}} = {{\bar B_{\rm{p}}^2}} / {{B_{\rm{a}}^2}}$          & normalized internal inductance   \\
    ${\beta _{\rm{p}}} = {{2{\mu _0}\bar p}} / {{B_{\rm{a}}^2}}$          & poloidal beta   \\
    $ W = \int_0^V {\frac{3}{2}p{\rm{d}}V'}$ & stored plasma energy \\
    $ q_0$, $q_{95}$ & safety factor at $\bar \psi = 0,\ 0.95$ \\
    \end{tabular}
\end{table}

### 2015-06-03 Results.tex
Response to review (first version, TO BE CHECKED)

\begin{figure*}[!htb]
\centering   %\begin{center}
\hfill{}
\includegraphics[width=18cm]{example_6962.pdf}
\hfill{}
%\end{center}
\caption{EFIT++ reconstructed pressure profiles and contours of $\bar\psi=\left(0.5,1\right)$ and VacTH LCFS from FREEBIE data, shot 6962 with Thomson scattering pressure profiles. EFIT++ parameters: $n_\mathrm{mp} = 16$, $n_\mathrm{fl} = 4$, $n_{FF'} = 1$. VacTH parameters: $n_\mathrm{mp} = 8$, $n_\mathrm{fl} = 16$, $n_P = n_Q = 4$. 3~\% random input data noise is used in the case of VacTH and EFIT++ with $n_{p'} = 2$, zero noise otherwise.}
\label{fig:ex6962}
\end{figure*}

We have selected five time slices from COMPAS shots 4275 and 6962 (i.e. 10 cases in total) for the analysis. These cases include circular, elongated and diverted plasmas with different currents.

@@

\begin{figure*}[!htb]
\centering   %\begin{center}
\hfill{}
\includegraphics[width=18cm]{RZstats.pdf}
\hfill{}
%\end{center}
\caption{Absolute errors in LCFS extents for convergent cases. EFIT++ results in the first row, VacTH in the second row. S1 denotes linear $p'$ and $FF'$ in EFIT++ as well as FREEBIE, s2 denotes TS pressure profiles in FREEBIE and $n_{p'} = n_{FF'} = 1$ in EFIT++, s3 denotes TS pressure profiles in FREEBIE and $n_{p'} = 2$ in EFIT++. Full lines show the means. VacTH parameters are $n_P=n_Q=4$, ``8 mp, 16 fl'' denotes $n_\mathrm{mp}=8$, $n_\mathrm{fl}=16$. Input error is calculated as an average of $I_{\rm{p}}$, magnetic probes and flux loops values. Zero input error data are scattered for a better visibility.}
\label{fig:RZstats}
\end{figure*}

\begin{figure*}[!htb]
\centering   %\begin{center}
\hfill{}
\includegraphics[width=18cm]{kinetic_stats_opt.pdf}
\hfill{}
%\end{center}
\caption{Internal plasma parameters relative errors for EFIT++ reconstructions using $n_{p'}=n_{FF'}=1$ in the top row and more optimized $n_{p'}$ in the bottom row. Zero input error data are scattered for a better visibility.}
\label{fig:kinetic_stats}
\end{figure*}

@@

In order to get a global overview of EFIT++ and VacTH reconstruction properties on COMPASS, we perform a scan over major code parameters and signal noise levels. In particular, $n_{p',FF'} = 1,2 $, $(n_\mathrm{mp}, n_\mathrm{fl}) = (16, 4), (64, 4), (8, 16)$, $\epsilon = 0, 0.02, 0.04, 0.06$, $n_{P,Q} = 4, 5, 6$. The same cases as above (time slices of shots 4275 and 6962) are used as target equilibria. For shot 6962, equilibria with TS pressure profiles and with linear $p'$ and $FF'$. This means there is 15 different target equilibria in total.

Absolute errors of the reconstructed LCFS extents for convergent cases from the scan are shown in Fig. \ref{fig:RZstats}. We can observer that the LCFS reconstructed with EFIT++ for target linear $p'$ and $FF'$ profiles (selection 1) are within 1~cm errors. There are, however, cases with up to 3 cm errors in $Z_{\rm{min}}$ for the more realistic TS pressure profiles (selection 2) if $n_{p'}=1$ is used. This error can be reduced by using $n_{p'}=2$. Input errors do not pose major difficulties for EFIT++.

@@

EFIT++ internal plasma parameters reconstruction results are shown in Fig. \ref{fig:kinetic_stats}. It shows that purely magnetic reconstruction with $n_{p'}=n_{FF'}=1$ introduces (except for $I_{\rm{p}}$) a systematic error for realistic pressure profiles, i.e. for plasmas that do not have the same profile parametrization. 
It is known that magnetic reconstruction with EFIT is difficult for small circular plasmas (without additional constraints, particularly the stored energy) \cite{efit1985}.
This suggests that using $n_{p'}=1$ for circular plasmas and $n_{p'}=2$ for elongated and diverted plasmas might lead to better results. This is demonstrated in the bottom row of Fig. \ref{fig:kinetic_stats}.
Reconstructions with such optimized parameters do not suffer from the systematic error; however, they generally increase the error bars for target equilibria with linear $p'$ and $FF'$, especially for $q_0$. It is also notable that $\delta l_{\rm{i}} \cong 0.1$ for all $n_{p'}=n_{FF'}=1$ reconstructions.

>>>>

\begin{figure*}[!htb]
\centering   %\begin{center}
\hfill{}
\includegraphics[width=17.5cm]{example_6962.pdf}
\hfill{}
%\end{center}
\caption{EFIT++ reconstructed pressure profiles and contours of $\bar\psi=\left(0.5,1\right)$ and VacTH LCFS from FREEBIE data, shot 6962 with Thomson scattering pressure profiles. EFIT++ parameters: $n_\mathrm{mp} = 16$, $n_\mathrm{fl} = 4$, $n_{FF'} = 1$. VacTH parameters: $n_\mathrm{mp} = 8$, $n_\mathrm{fl} = 16$, $n_P = n_Q = 4$. 3~\% random input data noise is used in the case of VacTH and EFIT++ with $n_{p'} = 2$, zero noise otherwise.}
\label{fig:ex6962}
\end{figure*}

We have selected five time slices from COMPAS shots 4275 and 6962 (i.e. 10 cases in total) for the analysis. These cases include circular, elliptical (limited) and diverted plasmas with different currents.

@@

\begin{figure*}[!htb]
\centering   %\begin{center}
\hfill{}
\includegraphics[width=17.5cm]{RZstats.pdf}
\hfill{}
%\end{center}
\caption{Absolute errors in LCFS extents for convergent cases. EFIT++ results in the first row, VacTH in the second row. S1 denotes linear $p'$ and $FF'$ in EFIT++ as well as FREEBIE, s2 denotes TS pressure profiles in FREEBIE and $n_{p'} = n_{FF'} = 1$ in EFIT++, s3 denotes TS pressure profiles in FREEBIE and $n_{p'} = 2$ in EFIT++. Full lines show the means. VacTH parameters are $n_P=n_Q=4$, ``8 mp, 16 fl'' denotes $n_\mathrm{mp}=8$, $n_\mathrm{fl}=16$. Input error is calculated as an average of $I_{\rm{p}}$, magnetic probes and flux loops values. 
% Zero input error data are scattered for a better visibility.
}
\label{fig:RZstats}
\end{figure*}

\begin{figure*}[!htb]
\centering   %\begin{center}
\hfill{}
\includegraphics[width=17.5cm]{kinetic_stats_opt.pdf}
\hfill{}
%\end{center}
\caption{Internal plasma parameters relative errors for EFIT++ reconstructions using $n_{p'}=n_{FF'}=1$ in the top row and more optimized $n_{p'}$ in the bottom row. Zero input error data are scattered for a better visibility.}
\label{fig:kinetic_stats}
\end{figure*}

@@

In order to get a global overview of EFIT++ and VacTH reconstruction properties on COMPASS, we perform a scan over major code parameters and signal noise levels. In particular, $n_{p',FF'} = 1,2 $, $(n_\mathrm{mp}, n_\mathrm{fl}) = (16, 4), (64, 4), (8, 16)$, $\epsilon = 0, 0.02, 0.04, 0.06$, $n_{P,Q} = 4, 5, 6$. The same cases as above (time slices of shots 4275 and 6962) are used as target equilibria. For shot 6962, equilibria with TS pressure profiles and with linear $p'$ and $FF'$ are used. This means there is 15 different target equilibria in total.

Absolute errors of the reconstructed LCFS extents for convergent cases from the scan are shown in Fig. \ref{fig:RZstats}. We can observe that the LCFS reconstructed with EFIT++ for target linear $p'$ and $FF'$ profiles (selection 1) are within 1~cm errors. There are, however, cases with up to 3 cm errors in $Z_{\rm{min}}$ for the more realistic TS pressure profiles (selection 2) if $n_{p'}=1$ is used. This error can be reduced by using $n_{p'}=2$. Input errors do not pose major difficulties for EFIT++.

@@

EFIT++ internal plasma parameters reconstruction results are shown in Fig. \ref{fig:kinetic_stats}. It shows that purely magnetic reconstruction with $n_{p'}=n_{FF'}=1$ introduces (except for $I_{\rm{p}}$) a systematic error for realistic pressure profiles, i.e. for plasmas that do not have the same profile parametrization. 
It is known that magnetic reconstruction with EFIT is difficult for small circular plasmas (without additional constraints, particularly the stored energy) \cite{efit1985}.
This suggests that using $n_{p'}=1$ for circular plasmas and $n_{p'}=2$ for elliptical and diverted plasmas might lead to better results. This is demonstrated in the bottom row of Fig. \ref{fig:kinetic_stats}.
Reconstructions with such optimized parameters do not suffer from the systematic error; however, they generally increase the error bars for target equilibria with linear $p'$ and $FF'$, especially for $q_0$. It is also notable that $\delta l_{\rm{i}} \cong 0.1$ for all $n_{p'}=n_{FF'}=1$ reconstructions.
